---
title: "Assignment 2 - R Markdown Presentation & Plotly"
author: "Tan Siu Ching"
date: "2025-04-24"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This presentation was created for the Peer-Graded Assignment of the Developing Data Products course.  

It uses **Plotly** for interactive visualizations based on air quality data, synthetically generated to mimic real-world urban conditions. 

A simulated dataset, including pollutants and weather parameters.  

The presentation includes the date and fulfills the project rubric.


## **About the Dataset**

This dataset was **synthetically generated** to reflect realistic urban air quality conditions. 

It includes pollutants such as CO, NOx, NO₂, O₃, SO₂, PM2.5, and PM10, as well as weather variables like temperature, humidity, wind speed, and pressure.

### Limitations:
- Synthetic, not from real sensors.
- Not suitable for actual policy use.

### Data Source:
From Kaggle. View the dataset:  
[https://www.kaggle.com/datasets/khushikyad001/air-quality-data]


## **Scatter Plot: PM2.5 vs PM10**

```{r load-data, message=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(plotly)

# Load CSV directly from GitHub raw link
url <- "https://raw.githubusercontent.com/SCTAN-Msia/My_Developing_Data_Products_Assignment2/main/AirQualityData.csv"
air_data <- read_csv(url)

# Categorize temperature into 3 groups
air_data$temp_group <- cut(
  air_data$Temperature,
  breaks = quantile(air_data$Temperature, probs = c(0, 0.33, 0.66, 1), na.rm = TRUE),
  labels = c("Low Temp", "Medium Temp", "High Temp"),
  include.lowest = TRUE
)

# Reshape data for multiple pollutant comparison
pollutant_data <- air_data %>%
  select(`PM2.5`, PM10, NO2) %>%
  pivot_longer(cols = everything(), names_to = "Pollutant", values_to = "Value")

plot_ly(data = air_data,
        x = ~`PM2.5`,
        y = ~PM10,
        type = 'scatter', mode = 'markers',
        marker = list(size = 10, color = ~`PM2.5`, colorscale = 'Viridis', showscale = TRUE)) %>%
  layout(title = "PM2.5 vs PM10 Air Pollution Levels",
         xaxis = list(title = "PM2.5 (µg/m³)"),
         yaxis = list(title = "PM10 (µg/m³)"))
```
## Box Plot: PM2.5 by Temperature Group
```{r}

plot_ly(data = air_data,
        x = ~temp_group,
        y = ~`PM2.5`,
        type = "box",
        color = ~temp_group,
        boxpoints = "none",
        jitter = 0.3,
        pointpos = -1.8,
        colors = "Set2") %>%
  layout(title = "PM2.5 Distribution by Temperature Group",
         xaxis = list(title = "Temperature Category"),
         yaxis = list(title = "PM2.5 (µg/m³)"))
```
## Box Plot by Comparing the Multiple Pollutants (PM2.5, PM10, and NO2)
```{r box-multipollutants-nodots, echo=FALSE}
plot_ly(data = pollutant_data,
        x = ~Pollutant,
        y = ~Value,
        type = "box",
        color = ~Pollutant,
        boxpoints = "none",  # ⛔ No individual dots
        colors = "Set1") %>%
  layout(title = "Pollutant Level Comparison: PM2.5, PM10, and NO2",
         xaxis = list(title = "Pollutant"),
         yaxis = list(title = "Concentration (µg/m³)"))
```